Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-23 articles/formalism/sections/kernel_discussion.tex
Minor edits through 3.5

The resultant set of radial function are depicted in \fig{fig:rad_ker_decay}. The amplitude of these radial function scales up as $\propto \ell_{\rm max}^2$, for clarity in the plot their global maxima are normalized to unity.  This normalization highlights the key feature, that on increasing $\ell_{\rm max}$ the radial kernels shift left, attaining their global maxima at progressively small angular distances $\beta$.  At intermediate values of $\beta$, the envelope of the radial functions is fit well by a power law $ \propto \beta^{-n}$.  As the band limit changes, the small angle portion of the radial functions have a similar shape.

These finding are neatly summarized in the observation that the radial functions with different maximum multipoles are approximately self-similar and related by this telescoping and scaling property:
$${}_rf(\beta,2,\ell_{\rm max}) \approx \Big[\frac{\ell_{\rm max}}{\ell'_{\rm max}}\Big]^2{}_rf(\beta'=\frac{\ell'_{\rm max}}{\ell_{\rm max}} \beta ,2,\ell'_{\rm max}) \qquad \revisit{\mbox{(small angles)}},$$
where ${}_rf , r \in [\mm, \md,\mi]$ denotes all the different radial functions. This telescoping property encapsulates both the amplitude scaling and leftward shift of the radial kernels on increasing the maximum multipole. \revisit{Specifically the global maxima of the kernel with maximum multipole $\ell_{max}$ is amplified(supressed) by a factor $(\ell_{\rm max}/\ell'_{\rm max})^2$ and the oscillating part is compressed(expanded) by a factor $(\ell'_{\rm max}/\ell_{\rm max})$ with respect to the kernel defined by the maximum multipole $\ell'_{\rm max}$ when $\ell_{\rm max} > \ell'_{\rm max}$($\ell_{\rm max} < \ell'_{\rm max}$).}

@@

The radial function ${}_\mm f(\beta)$ is oscillatory, regardless of the maximum $\ell_{\rm max}$. The envelope of the oscillation decays with angular separation at intermediate $\beta$ but increases as $\beta \rightarrow \pi$.  We find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.6}$ at intermediate $\beta$.  However, when we average over several oscillations with a moving window, we find that the average behavior at intermediate $\beta$ scales as $\beta^{-2}$, visible in \fig{fig:fbeta_avg}. We set the width of the smoothing window to $0.3\beta_0$, this results in the narrowing of the smoothing width as the band limit increases so as to keep the number of oscillations averaged over, remain roughly the same.  This agrees with the hypothesis for the \revisit{flat sky case} \comment{is this flat sky or continuum case?}, for which e.g. Zaldarriaga (2001) \cite{Zaldarriaga2001a} argued that form of the radial function has to be ${}_\mm f(\beta>0)\simeq\beta^{-2}$ to ensure that the Fourier modes of Stokes $Q/U$ relate to those of $E/B$ merely by rotations (with no scale dependence). In addition, we find that ${}_{\md}f$ under a moving average shows a similar scaling of $\beta^{-2}$, as seen in \fig{fig:dfbeta_avg}.
Applying a moving average on radial functions with increasing band limit, we see that a larger fraction of angular domain matches the scaling of $\beta^{-2}$.   Extrapolating this trend to the case of a very high band limit one expects this behavior to approach $\beta^{-2}$ across the domain $\beta\in[\epsilon,\pi-\epsilon]$ for $\epsilon \ll 1$. In essence the moving average can be understood as the averaging resulting from discrete sampling of a continuum field (one with an infinite band limit) where the width of the averaging window corresponds to the size of the pixel.

>>>>

The resultant set of radial function are depicted in \fig{fig:rad_ker_decay}. The amplitude of these radial function scales up as $\propto \ell_{\rm max}^2$.  For clarity in the plot, their global maxima are normalized to unity.  This normalization highlights the key feature, that on increasing $\ell_{\rm max}$ the radial kernels shift left, attaining their global maxima at progressively small angular distances $\beta$.  At intermediate values of $\beta$, the envelope of the radial functions is fit well by a power law $ \propto \beta^{-n}$.  As the band limit changes, the small angle portion of the radial functions have a similar shape.

These finding are neatly summarized in the observation that the radial functions with different maximum multipoles are approximately self-similar over many oscillations, and are related by this telescoping and scaling property:
\begin{equation}{}_rf(\beta,2,\ell_{\rm max}) \approx \Big[\frac{\ell_{\rm max}}{\ell'_{\rm max}}\Big]^2{}_rf(\beta'=\frac{\ell'_{\rm max}}{\ell_{\rm max}} \beta ,2,\ell'_{\rm max}),\end{equation}
where ${}_rf , r \in [\mm, \md,\mi]$ denotes all the different radial functions. This telescoping property encapsulates both the amplitude scaling and leftward shift of the radial kernels on increasing the maximum multipole. {Specifically, when $\ell_{\rm max} > \ell'_{\rm max}$, the global maxima of the kernel with maximum multipole $\ell_{\rm max}$ is amplified by a factor $(\ell_{\rm max}/\ell'_{\rm max})^2$ with respect to the kernel defined by the maximum multipole $\ell'_{\rm max}$, while the period of the oscillation is compressed by a factor $(\ell'_{\rm max}/\ell_{\rm max})$. The detailed accuracy of this approximation varies with the angular separation $\beta$ and the difference in the maximum multipoles ($\ell_{\rm max}$, $\ell'_{\rm max}$)  used in evaluating the two functions.}

@@

The radial function ${}_\mm f(\beta)$ is oscillatory, regardless of the maximum $\ell_{\rm max}$. The envelope of the oscillation decays with angular separation at intermediate $\beta$ but increases as $\beta \rightarrow \pi$.  We find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.6}$ at intermediate $\beta$.  However, when we average over several oscillations with a moving window, we find that the average behavior at intermediate $\beta$ scales as $\beta^{-2}$, visible in \fig{fig:fbeta_avg}. We set the width of the smoothing window to $0.3\beta_0$, and this results in the narrowing of the smoothing width as the band limit increases, so as to keep the number of oscillations averaged over roughly the same.  This agrees with the hypothesis for the {flat sky and continuum case}, for which e.g. Zaldarriaga (2001) \cite{Zaldarriaga2001a} argued that form of the radial function has to be ${}_\mm f(\beta>0)\simeq\beta^{-2}$ to ensure that the Fourier modes of Stokes $Q/U$ relate to those of $E/B$ merely by rotations (with no scale dependence). In addition, we find that ${}_{\md}f$ under a moving average shows a similar scaling of $\beta^{-2}$, as seen in \fig{fig:dfbeta_avg}.
Applying a moving average on radial functions with increasing band limit, we see that a larger fraction of angular domain matches the scaling of $\beta^{-2}$.   Extrapolating this trend to the case of a very high band limit one expects this behavior to approach $\beta^{-2}$ across the domain $\beta\in[\epsilon,\pi-\epsilon]$ for $\epsilon \ll 1$. In essence the moving average can be understood as the averaging resulting from discrete sampling of a continuum field (one with an infinite band limit) where the width of the averaging window corresponds to the size of the pixel.
